Query Optimization for Dynamic Imputation

### 2017-05-11 paper/experiments.tex

\begin{table*}
  \centering
  \input{queries}
\end{table*}

@@

\Cref{fig:runtimes} shows a summary of the performance results. The quality-optimized
queries are an \todo{update/validate: order-of-magnitude faster than imputing on the base tables. We can get
another order-of-magnitude speedup when optimizing for performance}. This performance
differential means it is feasible
to explore multiple imputations operations (including more expensive operators) when using
\ProjectName{}, in contrast to the traditional approach of base table imputation.

@@

\todo{update: In the median across a variety of queries and choices of $\alpha$, planning took
\planningmediantime{} ms, 
constituting \planningruntimepercent{} percent of total runtime. In all cases, the optimizer
returned a query plan within \planningmaxtime{} ms, with times similar across values of $\alpha$.}

@@

We explored the practical implications of our planning algorithm by constructing a series of queries with increasing number of joins.
We used the CDC tables and evaluated a combination of self-joins and joins across reference tables. Planning times for queries
involving 1 to 5 joins, which although limited is a practical value for real world exploratory queries, take less than a second.
\todo{update: For queries with 7 joins, planning time increases substantially to 29 seconds, on average. We introduced an optimization that reduces
the plan space by maintaining only approximations of the Pareto set for queries involving a large number of joins.
With this optimization, the same set of queries have an average planning time of 16 seconds (2 seconds SE).}

>>>>

\begin{table*}
  \centering
  \input{queries}
  \vspace{1ex}
  \caption{Queries used in our experiments.}\label{tbl:queries}
\end{table*}

@@

\Cref{fig:runtimes} shows a summary of the performance results. The quality-optimized
queries are an \todo{update/validate: order-of-magnitude faster than imputing on the base tables. We get another order-of-magnitude speedup when optimizing for performance}. This performance
differential means it is feasible
to explore multiple imputation operations (including more expensive operators) when using
\ProjectName{}, in contrast to the traditional approach of base table imputation.

@@

@@

\todo{Update: Planning times are reasonable, with a median planning time of \planningmediantime{} ms. over the benchmark queries, constituting \planningruntimepercent{}\% of total runtime.
In all cases, the optimizer returned a query plan within \planningmaxtime{} ms.}
The choice of $\alpha$ has no impact on planning times, because query selection is done after collecting the Pareto frontier.

However, \ProjectName{}'s planning algorithm is still exponential in the number of joins in the plan, and the use of Pareto sets exacerbates this exponential growth.
We explored the limits our planning algorithm by constructing a series of queries with increasing number of joins.
We used the CDC tables and evaluated a combination of self-joins and joins across reference tables.
Planning times for queries involving 1--5 joins, which is a practical value for real world exploratory queries, take less than a second.
\todo{update: For queries with 7 joins, planning time increases substantially to 29 seconds, on average. We introduced an optimization that reduces
the plan space by maintaining only approximations of the Pareto set for queries involving a large number of joins.
With this optimization, the same set of queries have an average planning time of 16 seconds (2 seconds SE).}

### 2017-05-11 paper/queries.tex
removed redundant filter in a query and addressed in rebuttal

\newcounter{queryno}
\begin{tabular}{cl}
\toprule
\# & \multicolumn{1}{c}{Queries on CDC data} \\
\midrule
1 & 
\begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT income, AVG(cuff_size) FROM demo, exams 
WHERE demo.id = exams.id AND height >= 150 GROUP BY income;
\end{lstlisting}
\end{minipage}\refstepcounter{queryno}\label[query]{q1} \\
2 & 
\begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT income, AVG(creatine) FROM demo, exams, labs 
WHERE demo.id = exams.id AND exams.id = labs.id AND income >= 13 AND income <= 15 AND weight >= 63 GROUP BY income;
\end{lstlisting}
\end{minipage}
\refstepcounter{queryno}\label[query]{q2} \\
3 & 
\begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT AVG(blood_lead) FROM demo, exams, labs 
WHERE demo.id = labs.id AND labs.id = exams.id AND age_yrs <= 6;
\end{lstlisting}
\end{minipage}\refstepcounter{queryno}\label[query]{q3}\\
4 & 
\begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT gender, AVG(blood_pressure_systolic) FROM demo, labs, exams 
WHERE demo.id = labs.id AND labs.id = exams.id AND body_mass_index >= 30 GROUP BY gender;
\end{lstlisting}
\end{minipage}\refstepcounter{queryno}\label[query]{q4}\\
5 & 
\begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT AVG(waist_circumference) FROM demo, exams 
WHERE demo.id = exams.id AND height >= 150 AND weight >= 100;
\end{lstlisting}
\end{minipage}\refstepcounter{queryno}\label[query]{q5}\\
\# & \multicolumn{1}{c}{Queries on freeCodeCamp data} \\
\midrule
6 & 
\begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT attendedbootcamp, AVG(income) FROM fcc
WHERE income >= 50000 GROUP BY attendedbootcamp;
\end{lstlisting}
\end{minipage}\refstepcounter{queryno} \label[query]{q6} \\
7 & 
\begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT AVG(commutetime) FROM fcc
WHERE gender = "female" AND countrycitizen = "United States";
\end{lstlisting}
\end{minipage}\refstepcounter{queryno} \label[query]{q7} \\
8 & 
\begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT schooldegree, AVG(studentdebtowe) FROM fcc
WHERE studentdebtowe > 0 AND schooldegree >= 0 GROUP BY schooldegree;
\end{lstlisting}
\end{minipage}\refstepcounter{queryno} \label[query]{q8}\\
9 & 
\begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT attendedbootcamp, AVG(gdp_per_capita) FROM fcc, gdp
WHERE fcc.countrycitizen = gdp.country AND age >= 18
GROUP BY attendedbootcamp;
\end{lstlisting}
\end{minipage}\refstepcounter{queryno} \label[query]{q9}\\
\bottomrule
\end{tabular}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

>>>>

\newcounter{queryno}
{
  \lstset{breaklines=true, basicstyle=\scriptsize\ttfamily}
\begin{tabular}{cl}
  \toprule
  \# & \multicolumn{1}{c}{Queries on CDC data} \\
  \midrule
  1 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines, basicstyle=\scriptsize\ttfamily]
SELECT income, AVG(cuff_size) FROM demo, exams 
WHERE demo.id = exams.id AND height >= 150 GROUP BY income;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno}\label[query]{q1} \\[-1.5ex]
  2 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines, basicstyle=\scriptsize\ttfamily]
SELECT income, AVG(creatine) FROM demo, exams, labs WHERE demo.id = exams.id AND exams.id = labs.id AND income >= 13 AND income <= 15 AND weight >= 63 GROUP BY income;
\end{lstlisting}
      \end{minipage}
      \refstepcounter{queryno}\label[query]{q2} \\[-1.5ex]
  3 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT AVG(blood_lead) FROM demo, exams, labs 
WHERE demo.id = labs.id AND labs.id = exams.id AND age_yrs <= 6;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno}\label[query]{q3}\\[-1.5ex]
  4 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT gender, AVG(blood_pressure_systolic) FROM demo, labs, exams 
WHERE demo.id = labs.id AND labs.id = exams.id AND body_mass_index >= 30 GROUP BY gender;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno}\label[query]{q4}\\[-1.5ex]
  5 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT AVG(waist_circumference) FROM demo, exams 
WHERE demo.id = exams.id AND height >= 150 AND weight >= 100;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno}\label[query]{q5}\\
  \midrule
  \# & \multicolumn{1}{c}{Queries on freeCodeCamp data} \\
  \midrule
  6 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT attendedbootcamp, AVG(income) FROM fcc WHERE income >= 50000 GROUP BY attendedbootcamp;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno} \label[query]{q6} \\[-1.5ex]
  7 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT AVG(commutetime) FROM fcc WHERE gender = "female" AND countrycitizen = "United States";
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno} \label[query]{q7} \\[-1.5ex]
  8 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT schooldegree, AVG(studentdebtowe) FROM fcc WHERE studentdebtowe > 0 GROUP BY schooldegree;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno} \label[query]{q8}\\[-1.5ex]
  9 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT attendedbootcamp, AVG(gdp_per_capita) FROM fcc, gdp 
WHERE fcc.countrycitizen = gdp.country AND age >= 18 GROUP BY attendedbootcamp;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno} \label[query]{q9}\\[-1.5ex]
  \bottomrule
\end{tabular}
}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
